Heuristics for the Score-Constrained Strip-Packing Problem

### 2018-04-24 PAPER/main.tex
moved theorem to end, list for rematching criteria, BPP NP-hard after SCSPP definition

\begin{definition}
	\label{defn:cpop}
	Let $\mathcal{M}$ be a multiset of unordered pairs of positive integers $\mathcal{M} = \{\{a_1, b_1\}, \{a_2,b_2\},...,\{a_n,b_n\}\}$, and let $\mathcal{T}$ be an ordering of the elements of $\mathcal{M}$ such that each element is a tuple. The Constrained Ordering Problem (COP) consists of finding a solution $\mathcal{T}$ such that, given a fixed value $\tau \in \mathbb{Z}^{+},$
	\begin{equation}
		\label{eqn:cpop}
		\textup{\textbf{rhs}}(i) + \textup{\textbf{lhs}}(i+1) \geq \tau \hspace{5mm} \forall \hspace{1mm} i \in \{1,2,..., n-1\},
	\end{equation}
	where \textup{\textbf{lhs}($i$)} and \textup{\textbf{rhs}($i$)} denote the left- and right-hand values of the $i$th tuple. The inequality is referred to as the \textup{vicinal sum constraint}.
\end{definition}

@@

The remainder of this article will firstly, in Section 2, detail a polynomial-time algorithm that is able to produce a solution to any instance of the COP, if a solution exists. Section 3 describes the heuristics that will be used to find feasible solutions to the SCSPP, and discusses the associated advantages and disadvantages. A comparison of these heuristics and an analysis of the results will be provided in Section 4, and finally Section 5 concludes the paper and proposes some potential directions for future work.

\section{Solving the CPOP}
\label{sec:ahca}

%We shall now present an algorithm for solving the COP, which operates by expressing the problem instance graphically. 
We shall now present an algorithm for solving the COP in polynomial time. The original(underlying) algorithm was created by \cite{becker2010}, and determines whether or not a solution exists for a given instance. We have extended this algorithm so that, if a solution does indeed exist, the algorithm is able to formulate and present us with this final solution. This is especially useful for problems such as the previously mentioned strip-packing problem, where a tangible solution is required in order to place the items in the correct order and orientation.

Let $\mathcal{M}$ be an instance of the COP of cardinality $n$ as described in Definition \ref{defn:cpop}. Then, a graph $G$ is defined by a vertex set $V = \{v_1, v_2, ...v_{2n+2}\}$, where each vertex is assigned a value from $\mathcal{M}$ such that $v_1 \leq v_2 \leq ... \leq v_{2n+2}$. Note that there are two extra ``dominating'' vertices, $v_{2n+1}$ and $v_{2n+2}$, which have values equivalent to $\tau$. In $\mathcal{M}$, the values that make up a pair are referred to as ``partners''. Thus, let $p: V \to V$ be a bijective function that associates each vertex $v_i \in V$ on $G$ with its partner, $v_j = p(v_i)$. $G$ comprises two distinct edge sets: $B$, which contains edges between vertices that are partners, and $R$, which consists of edges between vertices whose corresponding values in $\mathcal{M}$ can be arranged alongside one another. That is, for two distinct vertices $v_i, v_j$,

@@

provided $p(v_i) \neq v_j$. Hence, $G=(V, B \cup R)$ is an undirected graph with $2n + 2$ vertices and edge sets $B$ and $R$, such that $B \cap R = \emptyset$. Observe that $B$ is in fact a perfect matching, with $|B|= n+1$.

\begin{figure}	
\centering
\begin{subfigure}[h]{0.45\textwidth}
	\includestandalone[width=\textwidth]{figures/partners}
	\caption{\textcolor{red}{partners.}}	
	\label{fig:partners}
\end{subfigure} \hspace{40pt}
\begin{subfigure}[h]{0.45\textwidth}
	\includestandalone[width=\textwidth]{figures/threshold}
	\caption{\textcolor{red}{threshold.}}	
	\label{fig:threshold}
\end{subfigure}
\caption{$v_0 = 1, v_1 = 1, v_2 = 2, v_3 = 2, v_4 = 3, v_5 = 3, v_6 = 4, v_7 = 4, v_8 = 4, v_9 = 5, v_{10} = 6, v_{11} = 7, v_{12} = 7, v_{13} = 7$}
\end{figure}

@@

\begin{theorem}
	\label{thm:cpopsoln}
	There exists a feasible solution $\mathcal{T}$ to an instance $\mathcal{M}$ of the COP if and only if its corresponding graph $G(V, B\cup R)$ contains an alternating Hamiltonian cycle.
\end{theorem}
\begin{proof}
	Let $\mathcal{A} = \{v_1, v_2, ..., v_{2n+1}, v_{2n+2}\}$ be an alternating Hamiltonian cycle on $G$, where $v_{2n+1}$ and $v_{2n+2}$ are the dominating vertices. By definition, $\mathcal{A}$ contains all vertices on $G$, and all $n+1$ edgs in $B$. Thus, there exists a set of $n+1$ edges from $R$ that, together with the edges in $B$, form the alternating Hamiltonian cycle. Now, removing the dominating vertices and any incident edges gives rise to an alternating Hamiltonian path $\mathcal{A}' = \{v_1, v_2, ..., v_{2n-1}, v_{2n}\}$, whose first edge $(v_1, v_2)$ and last edge $(v_{2n-1}, v_{2n})$ are in $B$. The order of the values corresponding to each vertex of this path form a feasible solution $\mathcal{T}$, where values that are partners make up a tuple. \textcolor{OrangeRed}{edges R vicinal sum constraint equation 2}. \qed
\end{proof}

@@

During this algorithm, if a vertex $v_i$ is not adjacent to any other unmatched vertex except its partner $p(v_i)$, the preceeding vertex $v_{i-1}$ can be ``rematched'', provided $v_i$ is not the first vertex (i.e $i \neq 1$), $v_{i-1}$ has been matched, and $(v_{i-1}, p(v_i)) \in R$. Then, $v_{i-1}$ is matched with $p(v_i)$, and $v_i$ is matched with the vertex that was previously matched with $v_{i-1}$. It is guaranteed that MMCM will produce a maximum cardinality matching $R' \subseteq R$, and is able to do so in at most $O(n \log n)$ time.

@@

This algorithm, which will be referred to as the Alternating Hamiltonian Construction Algorithm (AHCA) is able to deduce whether a feasible solution $\mathcal{T}$ exists for any given instance $\mathcal{M}$ of the COP. If a solution does indeed exist, AHCA can produce this solution in at most $O(n^2)$ time.
%put theorem here
%explanation removing dominating vertices
%why is this different from becker

>>>>

\begin{definition}
	\label{defn:cop}
	Let $\mathcal{M}$ be a multiset of unordered pairs of positive integers $\mathcal{M} = \{\{a_1, b_1\}, \{a_2,b_2\},...,\{a_n,b_n\}\}$, and let $\mathcal{T}$ be an ordering of the elements of $\mathcal{M}$ such that each element is a tuple. The Constrained Ordering Problem (COP) consists of finding a solution $\mathcal{T}$ such that, given a fixed value $\tau \in \mathbb{Z}^{+},$
	\begin{equation}
		\label{eqn:cop}
		\textup{\textbf{rhs}}(i) + \textup{\textbf{lhs}}(i+1) \geq \tau \hspace{5mm} \forall \hspace{1mm} i \in \{1,2,..., n-1\},
	\end{equation}
	where \textup{\textbf{lhs}($i$)} and \textup{\textbf{rhs}($i$)} denote the left- and right-hand values of the $i$th tuple. The inequality is referred to as the \textup{vicinal sum constraint}.
\end{definition}

@@

The SCSPP is a generalisation of the classical one-dimensional bin-packing problem (BPP), where in the latter problem the minimum scoring distance $\tau$ can be said to be equal to zero. It follows that the SCSPP is at least as hard as the BPP, which is known to be NP-hard (\citealp{garey1979}), and so (under the assumption that $P \neq NP$) there is no known algorithm that is able to find an optimal solution for every instance of the SCSPP in polynomial time. Instead, heuristics can be used to find near-optimal solutions in a shorter amount of time.

The remainder of this article will firstly, in Section 2, detail a polynomial-time algorithm that is able to produce a solution to any instance of the COP, if a solution exists. Section 3 describes three heuristics that will be used to find feasible solutions to the SCSPP, and discusses the associated advantages and disadvantages. A comparison of these heuristics and an analysis of the results will be provided in Section 4, and finally Section 5 concludes the paper and proposes some potential directions for future work.

\section{Solving the COP}
\label{sec:ahca}

%We shall now present an algorithm for solving the COP, which operates by expressing the problem instance graphically. 
We shall now present an algorithm for solving the COP in polynomial time. The original(underlying) algorithm was created by \cite{becker2010}, and determines whether or not a solution exists for a given instance. We have extended this algorithm so that, if a solution does indeed exist, the algorithm is able to formulate and present us with this final solution. This is especially useful for problems such as the previously mentioned strip-packing problem, where a tangible solution is required in order to place the items in the correct order and orientation. \textcolor{OrangeRed}{USING GRAPH.}

\begin{itemize}
	\item Figure \ref{fig:partners} shows the graph with the values from the example above
	\item Model each of the values of all pairs in $\mathcal{M}$ as a vertex on a graph $G$
	\item Each vertex is assigned a value from $\mathcal{M}$
	\item Add an extra pair of vertices, dominating vertices, that are assigned values equal to $\tau$
	\item The values in $\mathcal{M}$, and thus the vertices on the graph, are arranged in pairs.
	\item This is represented on the graph by adding a set of ``blue'' edge, $B$, that are a perfect matching of all vertices on the graph
	\item Under such a matching, each vertex (each value in $\mathcal{M}$) has a partner that represents the other value in the pair.
	\item Then, we add to $G$ another set of ``red'' edges, $R$, that consists of edges between vertices whose corresponding values in $\mathcal{M}$ can be arrange alongside one another
	\item That is, for two distinct vertices, EQUATION, provided $p(v_i) \neq v_j$ i.e. not partners
	\item Bijective function $p: V \to V$ assoc each vertex with its partner, $v_j = p(v_i)$
	\item $G$ has $2n + 2$ vertices, edge sets $B$ and $R$ such that $B \cap R = \emptyset$
	\item $B = |n+1|$
	\item Dominating vertices are partners
	\item Dominating vertices will be removed at end, are not part of final solution
	\item Vertex set $V = \{v_1,...,v_{2n+2}\}$
	\item Each vertex assigned value from $\mathcal{M}$ s.t. $V-1 \leq ... \leq v_{2n+2}$
\end{itemize}

Let $\mathcal{M}$ be an instance of the COP of cardinality $n$ as described in Definition \ref{defn:cop}. Then, a graph $G$ is defined by a vertex set $V = \{v_1, v_2, ...v_{2n+2}\}$, where each vertex is assigned a value from $\mathcal{M}$ such that $v_1 \leq v_2 \leq ... \leq v_{2n+2}$. Note that there are two extra ``dominating'' vertices, $v_{2n+1}$ and $v_{2n+2}$, which have values equivalent to $\tau$. In $\mathcal{M}$, the values that make up a pair are referred to as ``partners''. Thus, let $p: V \to V$ be a bijective function that associates each vertex $v_i \in V$ on $G$ with its partner, $v_j = p(v_i)$. $G$ comprises two distinct edge sets: $B$, which contains edges between vertices that are partners, and $R$, which consists of edges between vertices whose corresponding values in $\mathcal{M}$ can be arranged alongside one another. That is, for two distinct vertices $v_i, v_j$,

@@

provided $p(v_i) \neq v_j$. Hence, $G=(V, B \cup R)$ is an undirected graph with $2n + 2$ vertices and edge sets $B$ and $R$, such that $B \cap R = \emptyset$. Observe that $B$ is also a perfect matching, with $|B|= n+1$.

\begin{figure}	
\centering
\begin{subfigure}[h]{0.45\textwidth}
	\includestandalone[width=\textwidth]{figures/partners}
	\caption{\textcolor{red}{partners.}}	
	\label{fig:partners}
\end{subfigure} \hspace{40pt}
\begin{subfigure}[h]{0.45\textwidth}
	\includestandalone[width=\textwidth]{figures/threshold}
	\caption{\textcolor{red}{threshold.}}	
	\label{fig:threshold}
\end{subfigure}
\caption{$v_1 = 1, v_2 = 1, v_3 = 2, v_4 = 2, v_5 = 3, v_6 = 3, v_7 = 4, v_8 = 4, v_9 = 4, v_{10} = 5, v_{11} = 6, v_{12} = 7, v_{13} = 7, v_{14} = 7$}
\end{figure}

@@

@@

During this algorithm, if a vertex $v_i$ is not adjacent to any other unmatched vertex except its partner $p(v_i)$, the preceeding vertex $v_{i-1}$ can be ``rematched'', provided that 
\begin{enumerate*}[label={(\alph*)}]
	\item $v_i$ is not the first vertex (i.e $i \neq 1$); 
	\item $v_{i-1}$ has been matched; and
	\item $(v_{i-1}, p(v_i)) \in R$.
\end{enumerate*}
Then, $v_{i-1}$ is matched with $p(v_i)$, and $v_i$ is matched with the vertex that was previously matched with $v_{i-1}$. It is guaranteed that MMCM will produce a maximum cardinality matching $R' \subseteq R$, and is able to do so in at most $O(n \log n)$ time. \textcolor{OrangeRed}{CITE.}

@@

This algorithm, which will be referred to as the Alternating Hamiltonian Construction Algorithm (AHCA) is able to deduce whether a feasible solution $\mathcal{T}$ exists for any given instance $\mathcal{M}$ of the COP. If a solution does indeed exist, AHCA can produce this solution in at most $O(n^2)$ time.

\begin{theorem}
	\label{thm:copsoln}
	There exists a feasible solution $\mathcal{T}$ to an instance $\mathcal{M}$ of the COP if and only if its corresponding graph $G(V, B\cup R)$ contains an alternating Hamiltonian cycle.
\end{theorem}
\begin{proof}
	Let $\mathcal{A} = \{v_1, v_2, ..., v_{2n+1}, v_{2n+2}\}$ be an alternating Hamiltonian cycle on $G$, where $v_{2n+1}$ and $v_{2n+2}$ are the dominating vertices. By definition, $\mathcal{A}$ contains all vertices on $G$, and all $n+1$ edgs in $B$. Thus, there exists a set of $n+1$ edges from $R$ that, together with the edges in $B$, form the alternating Hamiltonian cycle. Now, removing the dominating vertices and any incident edges gives rise to an alternating Hamiltonian path $\mathcal{A}' = \{v_1, v_2, ..., v_{2n-1}, v_{2n}\}$, whose first edge $(v_1, v_2)$ and last edge $(v_{2n-1}, v_{2n})$ are in $B$. The order of the values corresponding to each vertex of this path form a feasible solution $\mathcal{T}$, where values that are partners make up a tuple. \textcolor{OrangeRed}{edges R vicinal sum constraint equation 2}. \qed
\end{proof}
%put theorem here
%explanation removing dominating vertices
%why is this different from becker
